% !TEX TS-program = xelatex
\documentclass{exam}

\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{booktabs}

\firstpageheader{Econ 607 - Computational}{Assignment 3}{Page \thepage\ of \numpages}
\firstpageheadrule

\begin{document}
Assume a unit measure of firms living on $[0,1]$.  An individual firm faces problem of when to change prices.  Firms differ in the quality of their good which is given by $a_{i,t}$.  Assume that $a_{i,t}$ follows a growth rate process in logs
\begin{equation}
\log(a_{i,i})=\log(a_{i,t-1})+\epsilon_{i,t}
\end{equation} where $\epsilon_{i,t}$ is normally distributed with mean $0$, standard deviation $\sigma^2_\epsilon$ and is i.i.d. across agents and time.

Given their current price, $P_{i,t}$, firms face a demand for their good which is given by\footnote{For those who are curious this is the result of monopolistic competition and a Dixit-Stiglitz aggregator with elasticity of substitution given by $\gamma$.  We have normalized the aggregate price level to 1.}
\begin{equation}
D_{i,t}=a_{i,t}^{\gamma-1}\left(P_{i,t}\right)^{-\gamma}.
\end{equation}  We assume that the amount of labor required to produce good of quality $a_{i,t}$ is $a_{i,t}$.\footnote{Higher quality goods take more effort to produce.} If $\bar W$ is the wage rate in this economy, then the per-period profit of a firm is given by
\begin{equation*}
  	[P_{i,t}-a_{i,t}\bar W]D_{i,t}.
\end{equation*}

 Firms can change their prices $P_{i,t}$ every period but must pay a cost $\bar W^{1-\gamma}\kappa$ in order to do so.  Let $\beta$ be the discount factor of these firms. The firm's objective is to maximize
 \begin{equation}
 	\max_{P_{i,t}} \sum_{t=0}^\infty \beta^t\left([P_{i,t}-a_{i,t}\bar W]a_{i,t}^{\gamma-1}\left(P_{i,t}\right)^{-\gamma}-\bar W^{1-\gamma}\kappa1_{P_{i,t}\neq P_{i,t-1}}\right)
 \end{equation}given $P_{i,-1}$. $1_{P_{i,t}\neq P_{i,t-1}}$ is an indicator capturing if the firm changes its price.  
 \begin{enumerate}
 	\item Let $\mu_{i,t}=\frac{P_{i,t}}{a_{i,t}\bar W}$ be the markup (price over marginal cost) of firm $i$.  Show that you can write the firm's maximization problem as problem as 
 	\begin{equation}
 		\max_{\mu_{i,t}}\bar W^{1-\gamma}\sum_{t=0}^\infty \beta^t\left([\mu_{i,t}-1]\mu_{i,t}^{-\gamma}-\kappa1_{\mu_{i,t}\neq \frac{\mu_{i,t-1}}{\exp(\epsilon_{i,t})}}\right)
 	\end{equation}given $\mu_{i,-1}$
 	\item  Let $V(\mu\_)$ be the value of a firm entering with markup $\mu\_$, after it's current quality shock $\epsilon$ is realized,  which chooses not to change prices this period.  Let $W$ be the value of a firm which elects to change prices this period (why doesn't this depend on $\mu\_$?).  Write down the Bellman equation(s) of the firm.
 	\item  Let $\beta=0.9981$ (targeting a monthly discount factor), $\gamma =3$, $\kappa =0.01$ and $\sigma_\epsilon = 0.1$.  Write a program to solve the problem of the firm. (Use Gauss-Hermite quadrature to compute the expectation over $\epsilon_i$)
 	\item  When do firms change their prices? (You should find that firms change their prices once markups are outside of a certain range of the optimal markup)
 	\item  Write a program to simulate $100,000$ firms to compute an ergodic distribution of firm markups.  Adjust $\kappa$ and $\sigma_\epsilon$ so that in the ergodic distribution 21.9\% of firms change their price every month and the average size of price changes is $0.095$. Compute the average markup
 	\begin{equation*}
 		\left(\int_i\mu_{i,t}^{1-\gamma}\right)^\frac1{1-\gamma}
 	\end{equation*}in the steady state.
 	\item  Suppose that wage costs unexpectly increase by 5\% (this will reduce the markups of all firms proportionally by 5\%).  Starting from this new distribution simulate the path of of the average markup.  How long does it take firms to adjust prices to pass along these new costs?
 \end{enumerate}
\end{document} 
